---
title: "Meetup1"
author: "Catlin"
date: "September 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to handle unexpected values (1):  Pittsburgh bridges

Pittsburgh's three rivers, according to the documentation, are "A", "M", and "O".  

Note that the code below was written using "base R".  Later in the course, we'll look at a number of R pac

```{r bridges}

# sources: http://rpubs.com/robertwsellers/IS607_Homework2
# https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.names
library(RCurl)
if(!exists("df_bridges")) {
    URL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version2"
    x <- getURL(URL)
    df_bridges <- read.csv(textConnection(x), header=FALSE, sep=",")
}

colnames(df_bridges) <- c("IDENTIF", "RIVER", "LOCATION","ERECTED","PURPOSE","LENGTH","LANES","CLEAR-G","T-OR-D","MATERIAL","SPAN","REL-L","TYPE")
# bridges <- subset(df_bridges, PURPOSE == "HIGHWAY" & RIVER == "A", select = c(RIVER, ERECTED, TYPE))
bridges <- subset(df_bridges, select = c(RIVER, ERECTED, TYPE))
#View (bridges)

table(bridges$RIVER)


```

## How to handle unexpected values (2):  NYC airport weather

What does each weather observation represent?
How should we identify then handle outlier data?


```{r nycflights}
#install.packages("nycflights13")
library(nycflights13)
dim(weather)
#View(weather)
plot(weather$wind_speed)
subset(weather, wind_speed > 100, select = c(origin, year, month, day, hour, wind_speed))
```

