---
title: 'DATA 607:  Web APIs; Week 9'
author: "Walt Wells, Fall 2016"
output:
  html_document:
    theme: lumen
    toc: yes
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Our task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe. 

# Environment Prep
```{r, warning=FALSE, message=FALSE}
if (!require('jsonlite')) install.packages('jsonlite')
if (!require('dplyr')) install.packages('dplyr')
if (!require('stringr')) install.packages('stringr')
```

# Importing

Let's use the Top Stories V1 API and look for stories in the Technology section.  

## Key Config

We have stored our API developer Key in a file called 'configAPI.R' and sourced below.   The file is added to .gitignore.   To use the rest of the code below, uncomment and add your developer key below. 

```{r}
# get our API key
source("configAPI.R")

# others should uncomment and paste in their key below
# key <- "my_developer_key""

# add text required for call
addkey <- paste0("&api-key=", key)
```

## GET

```{r}
# add url and text required for call
url <- "https://api.nytimes.com/svc/topstories/v1/technology.json"
addurl <- paste0(url, "?")

# make call
feed <- fromJSON(paste0(addurl, addkey))
```

# Setup DF

```{r}
sub <- feed$results
sub <- sub %>% 
    filter(section=="Technology") %>% 
#    mutate(pub_date = str_extract(sub$published_date, "[0-9]{4}\\W[0-9]{2}\\W[0-9]{2}")) %>% 
    select(title, abstract, byline, url, published_date)
```

## Conclusion

```{r}
knitr::kable(sub)
```
